Template for indicator documentation

### 2023-09-13 DRAFT_NDVI_wetland.Rmd
ndvi wetland year regressions for good condition

### Regression analyses, NDVI across time
As a last step, we can investigate how NDVI has changed over time.

@@

# concatenate the three Satellite objects
allSatNDVI.wetland <- rbind(
SentinelNDVI.wetland[,!names(SentinelNDVI.wetland) %in% "subtype"],
ModisNDVI.wetland,
LandsatNDVI.wetland[,!names(LandsatNDVI.wetland) %in% "column_label"]
)
# add variable for Satellite identity
allSatNDVI.wetland$Sat <- c(
  rep("Sentinel",nrow(SentinelNDVI.wetland)),
  rep("Modis",nrow(ModisNDVI.wetland)),
  rep("Landsat",nrow(LandsatNDVI.wetland))
)
allSatNDVI.wetland$Sat <- factor(allSatNDVI.wetland$Sat,levels=c("Sentinel","Modis","Landsat"))
levels(allSatNDVI.wetland$Sat)
```
Now we can plot the time series for each main ecosystem type, and for each satellite separately  
```{r}
# plot 
allSatNDVI.wetland %>%
  ggplot( aes(x=year_jit, y=mean, color=Sat )) + 
  geom_point() +
  facet_wrap(~hovedtype, ncol = 1)
```
It is quite obvious that the NDVI values from the three satellites are not quantitatively comparable. They vary both in their placement along the y-axis and in their variance. Modis data are absent from the V4-ecosystem type because of a too large pixel size compared to the NiN-polygon sizes. In V10, Modis is poorly represented for the same reason.

As a final step, we model NDVI as a function of year and main ecosystem type separately for each Satellite
```{r}
model.wetland.time.Sent <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=SentinelNDVI.wetland)
model.wetland.time.Modi <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=ModisNDVI.wetland)
model.wetland.time.Land <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=LandsatNDVI.wetland)

>>>>

### NDVI across time - Sentinel, MODIS & LandSat
As a last step, we can investigate how NDVI has changed over time. For this we include data from the MODIS and LandSat satellites in addition to Sentinel.

@@

# concatenate the three Satellite objects
allSatNDVI.wetland <- rbind(
SentinelNDVI.wetland[,!names(SentinelNDVI.wetland) %in% "subtype"],
ModisNDVI.wetland,
LandsatNDVI.wetland[,!names(LandsatNDVI.wetland) %in% "column_label"]
)
# add variable for Satellite identity
allSatNDVI.wetland$Sat <- c(
  rep("Sentinel",nrow(SentinelNDVI.wetland)),
  rep("Modis",nrow(ModisNDVI.wetland)),
  rep("Landsat",nrow(LandsatNDVI.wetland))
)
allSatNDVI.wetland$Sat <- factor(allSatNDVI.wetland$Sat,levels=c("Sentinel","Modis","Landsat"))
levels(allSatNDVI.wetland$Sat)
```
Now we can plot the time series for each main ecosystem type, showing each satellite separately  
```{r}
# plot 
allSatNDVI.wetland %>%
  ggplot( aes(x=year_jit, y=mean, color=Sat )) + 
  geom_point() +
  facet_wrap(~hovedtype, ncol = 1)
```
It is quite obvious that the NDVI values from the three satellites are not quantitatively comparable. They vary both in their placement along the y-axis and in their variance. Modis data are absent from the V4-ecosystem type because of a too large pixel size compared to the NiN-polygon sizes. In V10, Modis is poorly represented for the same reason.

We can test if there is consistent tempral changes in NDVI by running regressions including year as a covariate.

We first investigate a model testing NDVI as a function of year and main ecosystem type for polygons in good condition, and we do so separately for each satellite.
```{r}
model.wetland.time.GodTilst.Sent <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=SentinelNDVI.wetland[SentinelNDVI.wetland$tilstand=="God",])
model.wetland.time.GodTilst.Modi <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=ModisNDVI.wetland[ModisNDVI.wetland$tilstand=="God",])
model.wetland.time.GodTilst.Land <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=LandsatNDVI.wetland[LandsatNDVI.wetland$tilstand=="God",])

summary(model.wetland.time.GodTilst.Sent)$coefficients$cond
summary(model.wetland.time.GodTilst.Modi)$coefficients$cond
summary(model.wetland.time.GodTilst.Land)$coefficients$cond
```

```{r}
model.wetland.time.Sent <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=SentinelNDVI.wetland)
model.wetland.time.Modi <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=ModisNDVI.wetland)
model.wetland.time.Land <- glmmTMB(mean_beta~year*hovedtype + (1|id), family=beta_family(),data=LandsatNDVI.wetland)
